A Beginners Guide to Perfomance Factors of Machine Learning Learning Applications with Focus on Neural Nets and Deep Learning

### 2018-01-09 paper.tex
modifications made

\section{Choice of Processing Units}
The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
\subsection{CPU}
As the first most naive option we have a look on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
\subsubsection{Single-thread}
The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required . But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Table I).
\subsubsection{Multi-thread}
A more complex approach consist of utilizing the multi-core characterisitics of modern CPUs, by splitting of training calculation with only requirements to already calculated values. This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm. The effect  continues to grow by using more threads but reaches it limits when using more threads than physical cores available.
\begin{table} 
\centering
\begin{tabular}{c c c c c}
\hline
Framework & 1 Thread & 2 Threads &4 Threads & 8 Threads\\\hline
Caffe & 1.324 & 0.790 & 0.578 & 15.444 \\
Tensorflow & 7.062 & 4.789 & 2.648 & 1.938 \\
Torch & 1.329 & 0.710 & 0.423 & na \\\hline
\end{tabular}
\caption{Training time(in s) for a mini-batch of size 64 of a Fully Connected Network(FCN) trained with synthetic data on a i7-3820 with 4-physical cores using 3 different commonly used Deep Learning frameworks\cite{shi2016benchmarking}}
\label{fig_ttfcn}
\end{table}
Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
\subsection{GPU}
A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs. For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}. Allmost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.

@@

%%data flow synchronization
Additional data flow is created when synchronizing GPUs with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations.

@@

To ease the effect of GPU to GPU communication,by relieveing the CPU of managing the , technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system.

>>>>

\section{Choice of Processing Units}
The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
\subsection{CPU}
As the first most naive option we have a look on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
\subsubsection{Single-thread}
The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required . But as can be thought of this does not uitilize the full capacity of the CPU and therefore results in performance loss in comparison to multithread usage(Table I).
\subsubsection{Multi-thread}
A more complex approach consist of utilizing the multi-core characterisitics of modern CPUs, by splitting of training calculation with only requirements to already calculated values. This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm. The effect  continues to grow by using more threads but reaches it limits when using more threads than physical cores available.
\begin{table} 
\centering
\begin{tabular}{c c c c c}
\hline
Framework & 1 Thread & 2 Threads &4 Threads & 8 Threads\\\hline
Caffe & 1.324 & 0.790 & 0.578 & 15.444 \\
Tensorflow & 7.062 & 4.789 & 2.648 & 1.938 \\
Torch & 1.329 & 0.710 & 0.423 & na \\\hline
\end{tabular}
\caption{Training time(in s) for a mini-batch of size 64 of a Fully Connected Network(FCN) trained with synthetic data on a i7-3820 with 4-physical cores using 3 different commonly used Deep Learning frameworks\cite{shi2016benchmarking}}
\label{fig_ttfcn}
\end{table}
Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
\subsection{GPU}
A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs. For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}. Almost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.

@@

%%data flow synchronization
Additional data flow is created when synchronizing GPUs with each other, by approaches described beforehand in "Choice of Processing Units", they each have their own advantage but can limit efficiency in certain situations. But both create an communication overhead because during transmission of values the GPU cannot operate on the neural network training.

@@

To ease the effect of GPU to GPU communication,by relieveing the CPU of managing the , technologies like GPUDirect from NVIDIA\cite{nvidiagpudirect2017} can be used allowing direct GPU to GPU communication if they are connected to the same PCIe bus, so this is only appplicable within a GPU Cluster or a single system and does not remove all communication overhead
